---
title: "Bandit Analysis"
output: github_document
---

Read in the data and store it as a data frame

```{r setup, include=FALSE}

library(rhdf5)
library(tidyverse)
library(ggplot2)

```


```{r helper functions}
# Load helper functions
setwd("banditCode")
source("Rcode/banditAnalysis.r")
source("Rcode/banditDataProcessing.r")
source("Rcode/banditArchetypalAnalysis.r")



```


```{r read data loop}
# Set the path to the hdf5 file
datafile <- 'dataset11092023.data'

# Load the hdf5 file
data <- h5dump(datafile,load=TRUE)

# Compute experiment stats
experimentData <- tabulateExperiments(data,datafile,2,'RawTelemteryDataLoessSmoothing')
# Save experment Data

save(experimentData,file=paste0("experimentData_",str_remove_all(datafile,".data"),".RData"))

#load(paste0("experimentData_",str_remove_all(datafile,".data"),".RData"))



```


```{r general data analysis}


# Designate the variables of interest from the data set
FeaturesOfIntrest <- c("P(0.25)",
                    "P(0.5)",
                    "P(1)",
                    "P(0.25->1)",
                    "P(1->0.5)",
                    "P(0.5->1)",
                    "P(Circle)")


# Retrieve the data set of interest
dataset <- experimentData[["SplitProbExperiments"]][["Split_1050_1800"]] #"Split_1050_1800"  "Split_1300_1800"

# Process the data set
dataset$Mouse <- as.character(dataset$Mouse)

# Make sure any column past column number 10 is numeric
dataset[,8:ncol(dataset)] <- sapply(dataset[,8:ncol(dataset)],as.numeric)

# Filter out any mice that have no trials (one drunk mice seems to have no trial past 5 minutes lol)
dataset <- dataset %>%
            filter(`NTrials` > 5)

dataset <-  dataset %>%
                 averageReplicates()

# Change States to consider name DTR+ as separate state  (DTRplus cuz some metaprogramming fucntionality is not working with the + sign)
dataset <- mutate(dataset, State = ifelse(MouseType == "DTR+", paste0(State,"DTRplus"), State))

# Filter Out GCAMP
dataset <- dataset %>%
            filter(MouseType != "GCAMP")

# Take only States that we are interested in (I think were completely dropping fed mice)
dataset <- dataset %>%
            filter(State %in% c("FastedEthDTRplus","FastedEth2DTRplus","FastedSalDTRplus",
                                "FastedVehDTRplus","FastedAlm","FastedEth",
                                "FastedEth2","FastedSal","FastedSnap","FastedVeh","Fasted"))



### NEEED TO FIX CV PLOTS BRRROOOOKKKEEENNNN
plotPCAMouse(dataset,c(FeaturesOfIntrest,"meanTimeToChoice","NTrials"),"LastHalfData",6,0.5,TRUE,c("RewardEvennes"))


```









```{r archtypal analysis plots}




# Define the adjency matrix for the predefined archetypal analysis
archMat3 <- matrix(c(1/3, 1/3, 1/3, 0.5, 0.5, 0.5, 1,  # Circle 
    	            0,   0.5, 0.5, 1,   1,   1,   0,  # GO
                    0.5,   0.5,   0,   0,   1,   0,   0), # Uncertainty seeking (maybe better to change to uncertainty avoiding  0.5,   0,   0.5,   1,   0,   1,   0)
                                                    nrow = 3,
                                                     ncol = 7, 
                                                     byrow = TRUE,
                                                    dimnames = list( c("Circle","GO","Uncertainty"),
                                                    FeaturesOfIntrest))

archMat4 <- matrix(c(1/3, 1/3, 1/3, 0.5, 0.5, 0.5, 1,  # Circle 
    	            0,   0.5, 0.5, 1,   1,   1,   0,  # GO
                    0.5,   0.5,   0,    0,   1,   0,   0, # Uncertainty seeking (maybe better to change to uncertainty avoiding  0.5,   0,   0.5,   1,   0,   1,   0)
                    0.5,   0,     0.5,  1,  0,   0.5,  0), # Certainty Seeking	
                                                    nrow = 4,
                                                     ncol = 7, 
                                                     byrow = TRUE,
                                                    dimnames = list( c("Circle","GO","Uncertainty","Certainty"),	
                                                    FeaturesOfIntrest))

# Designate the pairs of states to contrast in during plotting.
contrastList <- list(c("FastedAlm", "FastedVeh"),
                        c("FastedSnap", "FastedVeh"),
                        c("FastedEth","FastedSal"),
                        c("FastedEth2","FastedSal"),
                        c("FastedEthDTRplus","FastedSalDTRplus"),
                        c("FastedEth2DTRplus","FastedSalDTRplus")) 


# Do classical archetypal analysis (do plotting for 4 archetypes) This will generate some plots
AutoArchAnalysis <- autoArchetypesAnaPlots(dataset,"AutoFourArchetypes",4,c(FeaturesOfIntrest,"NTrials"),contrastList,FALSE) 

# Archatypal analysis with the 3 predefined archetypes
predefArchThree <- ArchAnalysisWithPredefined(archMat3,dataset,"PredefThree",FeaturesOfIntrest,contrastList,TRUE)
predefArchThree[["error"]]
predefThreePlots <- plotArchetypalAnalysisPredefined(TRUE,contrastList,predefArchThree[[1]],"PredefThree",archMat3)

# Archatypal analysis with 4 predefined archetypes
predefArchFour <- ArchAnalysisWithPredefined(archMat4,dataset,"PredefFour",FeaturesOfIntrest,contrastList,TRUE)
predefArchFour[["error"]]
predefFourPlots <- plotArchetypalAnalysisPredefined(FALSE,contrastList,predefArchFour[[1]],"PredefFour",archMat4)




```



```{r rolling estimate}



RollingEstimatesDataFrame <- gatherExperimentsRolling(experimentData[["RollingProbExperiments"]],c(FeaturesOfIntrest,
                                                                                "ngramEntropy_3",
                                                                                "ngramEntropy_4",
                                                                                "ngramEntropy_5",
                                                                                "meanTimeToChoice",
                                                                                "RewardEvennes",
                                                                                "totalDistance",
                                                                                "rewardPerDistance",
                                                                                "meanSpeed"))

# Plot line graphs of the rolling estimate of P(1) with ggplot
avargedData <- RollingEstimatesDataFrame %>%
                group_by(MidTimePoint,Mouse,State,MouseType,Sex) %>%
                summarise_if(is.numeric,mean) %>% ungroup() %>%
                mutate(State = ifelse(MouseType == "DTR+", paste0(State,"DTRplus"), State)) %>%
                filter(MouseType != "GCAMP") %>%
                filter(MidTimePoint > 400) %>%
                mutate(Mouse = as.character(Mouse)) %>%
                as.data.frame()

getpval <- function (modelobject) {
        f <- modelobject$fstatistic
        p <- pf(f[1],f[2],f[3],lower.tail=F)
        attributes(p) <- NULL
        return(p)
}

plotDifferencesTimeProgressionComparison <- function(avargedData,comparisonVec,varibleToCompare,timesegments=1) {

                datas <- filter(avargedData, State %in% comparisonVec)
                datas$State <- factor(datas$State,levels = comparisonVec)

                if(timesegments == 1) {
                        datas <- datas %>% mutate(TimeSegment = "Entire Probabilistic Phase") %>%
                                mutate(Duped = FALSE)
                } else {
                        datas <- datas %>%
                                mutate(TimeSegment = cut(MidTimePoint, breaks = timesegments,label = FALSE)) %>%
                                mutate(Duped = FALSE)

                        # Sort the data frame by the val column
                        for(i in 2:timesegments) {
                                for(mouse in unique(datas$Mouse)) {
                                        for(state in unique(datas$State)) {
                                                temp <- filter(datas, Mouse == mouse & State == state & TimeSegment == i)
                                                addrow <- temp[which.min(temp$MidTimePoint),]
                                                addrow$TimeSegment <- i-1
                                                addrow$Duped <- TRUE
                                                datas <- rbind(datas,addrow)
                                        }
                                }
                        }
                }


                forumlaStringNoState <- paste0("`",varibleToCompare,"`", "~ MidTimePoint")
                forumlaString <- paste0("`",varibleToCompare,"`", "~ MidTimePoint + State")
                forumlaStringWInteraction <- paste0("`",varibleToCompare,"`", "~ MidTimePoint + State + MidTimePoint:State")
                forumlaStringOnlyInteraction <- paste0("`",varibleToCompare,"`", "~ MidTimePoint + MidTimePoint:State")
        

                lmCoeffsDF <- data.frame(rowNames = c("Intercept ",
                                        "Time",
                                        "State",
                                        "Model (R^2)",
                                        "Intercept",
                                        "Time",
                                        "State",
                                        "Time:State",
                                        "Model (R^2)",
                                        "Intercept",
                                        "Time",
                                        "Time:State", 
                                        "Model (R^2)"))

                if(timesegments == 1) {
                        newStrings <- c("Entire Probabilistic Phase")
                } else {
                        ditt <- cut(datas$MidTimePoint,timesegments,dig.lab=5) %>%
                                                levels() %>%
                                                str_remove_all("\\(|\\)|\\]|\\[") %>%
                                                str_split( ",") %>%
                                                unlist() %>%
                                                as.numeric() %>%
                                                round(-1) 
                        newStrings <- c()
                        
                        for(i in 1:timesegments) {
                                symb <- ifelse(i == 1,"[","(")
                                newStrings <- c(newStrings,paste0(symb,ditt[(2*i)-1],",",ditt[2*(i)],"]"))
                        }
  
                }
                
                lmCompare <- list()

                for(slice in unique(datas$TimeSegment)) {
                        if(timesegments == 1) {
                                datasSlice <- filter(datas,Duped == FALSE)
                        } else {
                                datasSlice <- filter(datas, TimeSegment == slice)
                        }

                        datasFirst <- filter(datasSlice, State == comparisonVec[1])	
                        datasSecond <- filter(datasSlice, State == comparisonVec[2])
                        
                        firstStatelm <- lm(forumlaStringNoState,datasFirst)
                        secondStatelm <- lm(forumlaStringNoState,datasSecond)

                        differenceInIntercepts <- coef(secondStatelm)[1] - coef(firstStatelm)[1]
                        ratioOfCoeffcients <- coef(secondStatelm)[2]/coef(firstStatelm)[2]

                        lmCompare[[slice]] <- list(differenceInIntercepts,ratioOfCoeffcients)

                        comblm <-summary(lm(forumlaString,datasSlice))
                        comblmWInteraction <-summary(lm(forumlaStringWInteraction,datasSlice))
                        comblmOnlyInteraction <-summary(lm(forumlaStringOnlyInteraction,datasSlice))

                        lmCoeffsDF <- cbind(lmCoeffsDF,data.frame("Coefficients" = c(
                                        comblm$coefficients[,1],
                                        comblm$adj.r.squared,
                                        comblmWInteraction$coefficients[,1],
                                        comblmWInteraction$adj.r.squared,
                                        comblmOnlyInteraction$coefficients[,1],
                                        comblmOnlyInteraction$adj.r.squared
                                        ),
                                        "P-Value" = c(
                                        comblm$coefficients[,4],
                                        getpval(comblm),
                                        comblmWInteraction$coefficients[,4],
                                        getpval(comblmWInteraction),
                                        comblmOnlyInteraction$coefficients[,4],
                                        getpval(comblmOnlyInteraction)
                                        )))

                }
                
                colnames(lmCoeffsDF) <- 1:ncol(lmCoeffsDF)
                 
                colGroups <- c(1,rep(2,timesegments))
                colGroups <- setNames(colGroups, c(" ",newStrings))

                table <- lmCoeffsDF %>% dplyr::mutate_if(is.numeric, funs(as.character(signif(., 3)))) %>%
                kableExtra::kbl(
                                col.names = c(" ",rep(c("Coefficient","P-Value"),timesegments))) %>% 
                        kableExtra::pack_rows(paste0(varibleToCompare, " ~ Time + State"), 1, 4) %>%
                        kableExtra::pack_rows(paste0(varibleToCompare, " ~ Time + State + Time:State"), 5,9) %>%
                        kableExtra::pack_rows(paste0(varibleToCompare, " ~ Time + Time:State"), 10, 13) %>%
                        kableExtra::add_header_above(colGroups) %>%
                        kableExtra::add_header_above(c(" "=1,"Time Intervals (s)" = length(colnames(lmCoeffsDF))-1)) %>%
                        kableExtra::column_spec(1, bold = T, border_right = T) %>%
                        kableExtra::row_spec(c(4,9,13), underline = T) %>%
                        kableExtra::kable_classic_2(c("striped","condensed"),full_width = F)


                
                

                subtitleString <- c()
                for(i in 1:length(lmCompare)) {
                        subtitleString <- c(subtitleString,
                                             "Segment: ",newStrings[i],"\n",
                                             "    Difference in intercept = ",signif(lmCompare[[i]][[1]],3), "\n",
                                             "    Ratio of Slopes               = ",signif(lmCompare[[i]][[2]],3), "\n")
                }
                subtitleString <- paste0(subtitleString,collapse = "")

                p <- ggplot() +
                geom_line(data = datas,
                                aes_string(x = "MidTimePoint", y = paste0("`",varibleToCompare,"`"), group = "Mouse"),
                                alpha = 0.1,
                                size=1) +
                geom_line(data = datas,stat="smooth",
                        aes_string(x = "MidTimePoint", y = paste0("`",varibleToCompare,"`"), group = "TimeSegment"),
                        color = "#ff0000",
                        method = "lm",
                        se = TRUE,
                        size = 2) +
                geom_smooth(data =datas,
                        aes_string(x = "MidTimePoint", y = paste0("`",varibleToCompare,"`"), group = "TimeSegment"),
                        color = "#ff0000",
                        method = "lm",
                        linetype="dashed",
                        se = TRUE,
                        size = 0.01) +
                geom_line(data = datas, stat="smooth",
                        aes_string(x = "MidTimePoint", y = paste0("`",varibleToCompare,"`")),
                        color = "#001aff",
                        method = "loess",
                        linetype="dashed",
                        alpha=0.8,
                        se = FALSE,
                        size = 1) +  
                theme_bw() + 
                facet_wrap(~State, ncol = 2) +
                ggpubr::stat_regline_equation(data = datas,
                                                mapping = aes_string(x = "MidTimePoint", y = paste0("`",varibleToCompare,"`"),group="TimeSegment")) + 
                ggpubr::stat_cor(data = datas,
                                                mapping = aes_string(x = "MidTimePoint", y = paste0("`",varibleToCompare,"`"),group="TimeSegment"),label.x.npc = "middle") + 
                xlab("Time (s)") +
                guides(color = "none") + 
                ggtitle(paste0(varibleToCompare, " : ", comparisonVec[1]," vs. ",comparisonVec[2]),
                        subtitle = subtitleString)
                return(p)



}

plotFolder <- "plots/ComaparingStatesWithArchs"

timerSeriesArch <- ArchAnalysisWithPredefined(archMat4,avargedData,"TimeSeriesDataFourArchs",FeaturesOfIntrest,contrastList,TRUE)

avargedDataWarchalphas <- cbind(timerSeriesArch[[1]][,rownames(archMat4)],avargedData) 

featuresToCompare <- c(FeaturesOfIntrest,
                        rownames(archMat4),
                        "NTrials",
                        "ngramEntropy_3",
                        "ngramEntropy_4",
                        "ngramEntropy_5",
                        "meanTimeToChoice",
                        "RewardEvennes",
                        "totalDistance",
                        "rewardPerDistance",
                        "meanSpeed")

contrastList <- list(   c("Fasted", "FastedVeh"),
                        c("FastedVeh", "FastedAlm"),
                        c("FastedVeh", "FastedSnap"),
                        c("FastedSal","FastedEth"),
                        c("FastedSal","FastedEth2"),
                        c("FastedSalDTRplus","FastedEthDTRplus"),
                        c("FastedSalDTRplus","FastedEth2DTRplus")) 



for(i in 1:length(featuresToCompare)) {
        for(j in 1:length(contrastList)) {
                a <- plotDifferencesTimeProgressionComparison(avargedDataWarchalphas,contrastList[[j]],featuresToCompare[i],1)
                ggsave(paste0(plotFolder,"/",featuresToCompare[i],"_",contrastList[[j]][1],"_",contrastList[[j]][2],".png"),a,width = 10,height = 7)
        }
}


``` 


```{r archtypal analysis plots}
# Plot pca of archetypes

pcaArchs <- prcomp(predefArchFour[[1]][,1:4],center = TRUE,scale. = TRUE)

ggplot(cbind(pcaArchs$x,predefArchFour[[1]][,6:8]),aes(x = PC1, y = PC2, color = State)) +
        geom_point() +
        theme_bw() + 
        guides(color = guide_legend(override.aes = list(label = NULL)))


```

```{r generate plots}

# Create a function for plotting scatterplots with corr 

scatterWrapper <- function(archData,archtypesNames,variableOFInterest,name) {
    tempDF <- tidyr::pivot_longer(archData,cols = archtypesNames,names_to = "Archs",values_to = "Value")
    plot <- ggpubr::ggscatter(tempDF,x = "Value", y = variableOFInterest,add = "reg.line",conf.int = T,cor.coef = T) +
            facet_grid(~Archs ,scales = "free_x") +
            ggtitle(name) + 
            theme_bw() 

    return(plot)
}

dataArchComb <- cbind(predefArchFour[[1]],dataset)

scatterWrapper(dataArchComb,c("Circle","Certainty","Uncertainty","GO"),"rewardPerDistance", "Reward per Distance")

scatterWrapper(dataArchComb,c("P(0.25)","P(0.5)","P(1)","P(Circle)","P(DirectionChange)"),"rewardPerDistance", "Reward per Distance")

scatterWrapper(dataArchComb,c("Circle","Certainty","Uncertainty","GO"),"totalDistance", "Total Distance")

scatterWrapper(dataArchComb,c("P(0.25)","P(0.5)","P(1)","P(Circle)"),"totalDistance", "Total Distance")

scatterWrapper(dataArchComb,c("Circle","Certainty","Uncertainty","GO"),"RewardEvennes", "Reward Evennes")

scatterWrapper(dataArchComb,c("Circle","Certainty","Uncertainty","GO"),"meanTimeToChoice", "Mean time to choice")

ggpubr::ggscatter(dataset,x = "NTrials", y = "totalDistance",add = "reg.line",conf.int = T,cor.coef = T)



reps<- predefFourPlots[["representations"]]

tempDF <- reps[[3]]

tempDF$rewardPerDistance <- dataset$rewardPerDistance
	
ggplot() +
        geom_point(data = tempDF,aes(x = PC1, y = PC2, color =dataset$rewardPerDistance)) +
        scale_color_gradient(low = "lightgray", high = "red") +
        geom_text(data = reps[[4]],aes(x = PC1, y = PC2, label = Archs)) + 
        theme_bw() + 
        guides(color = guide_legend(override.aes = list(label = NULL))) 


# Get the column that has the highest value for each row
MaxArchDF <- predefArchFour[[1]][,1:7] %>%
            mutate(Archs = apply(predefArchFour[[1]][,1:4],1,function(x) colnames(predefArchFour[[1]][,1:4])[which(x == max(x))]))

MaxCounts <- MaxArchDF %>% group_by(State,Archs) %>% summarise(maxCounts = n())

ggplot(asd,aes(x = Archs, y = n, fill = Archs)) +
        geom_bar(stat = "identity",position = "dodge") +
        theme_bw() + 
        facet_wrap(~State, ncol = 3 ) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
        guides(fill = guide_legend(override.aes = list(label = NULL)))

```

 




```{r helper functions}
# Load helper functions
source("banditAnalysis.r")
source("banditDataProcessing.r")
source("banditArchetypalAnalysis.r")

```

```{r addd}

t.test(Circle ~ State,data = dataArchComb[which(dataArchComb$State %in% c("FastedVeh","FastedAlm")),])
t.test(GO ~ State,data = dataArchComb[which(dataArchComb$State %in% c("FastedVeh","FastedAlm")),])
t.test(Uncertainty ~ State,data = dataArchComb[which(dataArchComb$State %in% c("FastedVeh","FastedAlm")),])
t.test(Certainty ~ State,data = dataArchComb[which(dataArchComb$State %in% c("FastedVeh","FastedAlm")),])



 #### CREATE A LINEAR MODEL TO PREDICT THINGS FROM ARCHTYPAL VALS
```



filter(RollingEstimatesDataFrame,State == "FastedVeh") %>%
ggplot(aes(x = MidTimePoint, y = `P(1)`, color = Day)) +
        geom_line(alpha = 0.7) +
        geom_smooth(data = avargedData,
                        aes(x = MidTimePoint, y = `P(1)`),
                        alpha=0.4,
                        color = "#ff0000",
                        method = "lm",
                        linetype="dashed"
                        ) +
        theme_classic() + 
        facet_wrap(~Mouse, ncol = 4) +
        guides(color = "none")

ggplot(avargedData) +
        geom_smooth(aes(x = MidTimePoint, y = `P(1)`,color = as.character(Mouse),alpha=0.3),
                        alpha=0.4,
                        method = "lm",
                        se = FALSE,
                        linetype="dotted"
                        ) +
        geom_smooth(aes(x = MidTimePoint, y = `P(1)`,color = as.character(Mouse),alpha=0.3),
                        alpha=0.4,
                        method = "loess",
                        se = FALSE,
                        ) +
        theme_classic() + 
        xlab("Time (s)") +
        guides(color = "none")

####################
